---
title: "OpenHDF5"
author: "Kristin Braziunas"
date: "June 20, 2016"
output: html_document
---

## Learning to open HDF5 files in R

```{r load-libraries }

# load libraries
library(raster)
library(rhdf5)
library(rgdal)

```

## Let's open a file

```{r select-file }

# set file path
f <- "../NEONdata/D17-California/TEAK/2013/spectrometer/reflectance/Subset3NIS1_20130614_100459_atmcor.h5"

# view h5 structure
h5ls(f)

```

## Import spatial information

```{r import-spatial-info }

# import spatial info
mapInfo <- h5read(f, 
                  "map info",
                  read.attributes = TRUE)
mapInfo

```

## Grab reflectance metadata

```{r get-reflectance-metadata }

# read in reflectance data attributes
reflInfo <- h5readAttributes(f, "Reflectance")
reflInfo

# define scale factor
scaleFactor <- reflInfo$`Scale Factor`

# define no data value
noDataValue <- reflInfo$`data ignore value`  # it's a character
noDataValue <- as.numeric(noDataValue)
str(noDataValue)  # looks good

```

## Import data dimensions

```{r import-dims }

# open file for viewing
fid <- H5Fopen(f)

# open the reflectance dataset
did <- H5Dopen(fid, "Reflectance")
did  # note dimensions are in columns x rows x bands

# grab dataset dimensions
sid <- H5Dget_space(did)
dims <- H5Sget_simple_extent_dims(sid)$size
dims  # columns, rows, bands; r reads rows 1st, columns 2nd

# close all open connections
H5Sclose(sid)
H5Dclose(did)
H5Fclose(fid)

```

## Read in reflectance data

```{r read-data, results="hide" }

# extract slice of H5 file
b56 <- h5read(f,
              "Reflectance",
              index = list(1:dims[1], 1:dims[2], 56))
b56

class(b56)

```

## Convert data to matrix

```{r convert-matrix }

# convert to matrix
b56 <- b56[,,1]  # z dimension goes away, set to 1
class(b56)

# let's plot some data
image(b56)

# stretch the image by log transforming
image(log(b56), main = "Log transformed data")

# force non-scientific notation
options("scipen"=100, "digits"=4)

# look at histograms
hist(b56)

# histogram of stretched image
hist(log(b56))

```

## Time to clean up our data

Note that data is stored as integers rather than floats. This way, data takes up less space. Can use scale factor to convert to decimal point at later time.

```{r data-clean-up }

# assign no data values to object
b56[b56 == noDataValue] <- NA

# apply scale factor
b56 <- b56 / scaleFactor  # grabbed scale factor earlier
hist(b56)  # right skew

```

## Transpose data

```{r transpose-data }

# transpose data, flip rows and columns
b56 <- t(b56)
image(log(b56))  # still not quite right

```

## Create spatial extent

```{r create-spatial-extent }

# split out map info object
mapInfo <- strsplit(mapInfo, ",")
mapInfo <- unlist(mapInfo)  # remove nesting
mapInfo

# value at 3 element in list
mapInfo[3]  # note that it's a character

# define upper left hand corner coordinate
xMin <- as.numeric(mapInfo[4])
yMax <- as.numeric(mapInfo[5])

# get spatial resolution
xRes <- as.numeric(mapInfo[6])
yRes <- as.numeric(mapInfo[7])

# define lower right hand corner coordinate
xMax <- xMin + (dims[1] * xRes)
yMin <- yMax - (dims[2] * yRes)

# create extent object
rasExt <- extent(xMin, xMax, yMin, yMax)
rasExt

# create raster object
b56r <- raster(b56, 
               crs = CRS("+init=epsg:32611"))

extent(b56r) <- rasExt
b56r

# plot data
plot(b56r, main="Spatially referenced data")

```

## Import NEON functions

```{r import-using-functions }

# install devtools
# install.packages("devtools")
library(devtools)

# install_github("lwasser/neon-aop-package/neonAOP")
library(neonAOP)

b55 <- open_band(f,
                 bandNum = 55,
                 epsg = 32611)
plot(b55)

# import several bands
bands <- c(58, 34, 19)  # decreasing numeric order intentional to make r, g, b
bands

# create raster stack
RGBStack <- create_stack(f,
                         bands = bands,
                         epsg = 32611)
plot(RGBStack)  # plot 3 bands separately

# plot RGB image
plotRGB(RGBStack,
        stretch = "lin")  # make sure ordered correctly

# cir image
bands <- c(90, 34, 19)

# create raster stack
CIRStack <- create_stack(f,
                         bands = bands,
                         epsg = 32611)

# plot RGB image
plotRGB(CIRStack,
        stretch = "lin")  # make sure ordered correctly


```

